---
title: "Table Structure"
---

# Optimization

## Table Structure

- longer, narrower tables are generally better than wide, short tables
- tradeoff between query speed and index size
- NULLs
- [3NF](http://en.wikipedia.org/wiki/Third_normal_form)
    - Idea: don't duplicate data, use keys and more tables; keep like with like

## Keys & Indices

- PRIMARY: every row should have a *unique* identifier
    - a table may have only one PRIMARY key
    - generally bad practice not to have a PRIMARY KEY
- [INDEX]() (see below)
- UNIQUE: unlike PRIMARY, can have NULLs
- search may be $O(\log{n})$ instead of $O(n)$
- Optimizing queries with [EXPLAIN](http://www.postgresql.org/docs/9.4/static/sql-explain.html)

## Foreign Keys

- [Referential Integrity](https://en.wikipedia.org/wiki/Referential_integrity):
    - every value in a FOREIGN KEY must correspond to an entry in another table's PRIMARY KEY
    - unidirectional
    - one-to-one, one-to-many
- [Entity Relationship Schema](https://en.wikipedia.org/wiki/Entity%E2%80%93relationship_model)


## Partitioning

- [Inheritance](http://www.postgresql.org/docs/9.4/static/ddl-inherit.html)
- split the data into separate [logical blocks](http://www.postgresql.org/docs/9.4/static/ddl-partitioning.html)

## Housekeeping

- [`VACUUM ANALYZE`](http://www.postgresql.org/docs/9.4/static/sql-vacuum.html)

## Example
```{psql, eval=FALSE}
DROP INDEX IF EXISTS individuals_state_idx;
DROP INDEX IF EXISTS committees_cmte_st_idx;

-- wicked slow, running time forever...
-- note sequential scans
EXPLAIN SELECT sum(1) as N FROM committees c
JOIN individuals i ON c.cmte_st = i.state
WHERE i.state = 'MA';

-- Build index on big table
CREATE INDEX ON individuals (state);

-- now it uses the index, running time 
EXPLAIN SELECT sum(1) as N FROM committees c
JOIN individuals i ON c.cmte_st = i.state
WHERE i.state = 'MA';

-- Build index on small table
CREATE INDEX ON committees (cmte_st);

-- now it uses BOTH indexes, running time 
EXPLAIN SELECT sum(1) as N FROM committees c
JOIN individuals i ON c.cmte_st = i.state
WHERE i.state = 'MA';
```

# Activity

```{r, echo=FALSE, message=FALSE} 
require(knitr)
opts_chunk$set(echo = FALSE, eval=FALSE)
```

```{r connect}
require(RPostgreSQL)
con <- dbConnect(PostgreSQL(), host="localhost", user="postgres", password="postgres", dbname="airlines")
```

## Questions

1. An entity relationship schema shows how _____are related to each other in a database.
2. A primary key ____-ly identifies an observation in a ____.
3. A primary key in one table can be a foreign key in another table. Can the foreign key
alone uniquely describe a row of data in a table?
3. What's the purpose of primary keys and foreign keys?
4. Can primary keys have repeat values in a table?
5. Can foreign keys have repeat values in a table?
6. Are primary and foreign keys related only through one-to-one
relationships?

